---
title: "maps"
author: "hannah"
format: html
editor: visual
---

Eventually: a function that allows users to select:

-   An overlay variable describing HFC recipient organizations (at the ZIP code level)

-   A background variable describing the food environment in Los Angeles (at the census tract level)

The goal is to quickly visualize relevant data so that HoFoCo can assess:

-   Which geographies may be worth expanding into

-   Which areas may demonstrate the greatest need

**Overlay Variables (HFC Recipient Orgs, ZIP Code Level):**

-    `AvgTotMonthly`: Total number served monthly by each org, divided by the number of ZIP codes served by that org, then aggregated by ZIP code.

-   `AvgUniqueInds`: Unique individuals served monthly by each org, divided by the number of ZIP codes served by that org, then aggregated by ZIP code.

-   `HFC_count`: The number of HFC recipient organizations operating in each ZIP code.

**Background Variables (LA County, Census Tract Level)**

-   `LILATracts_halfAnd10`: Flag for low-income and low access when considering low accessibilty at 1/2 and 10 miles - from USDA Food Atlas

-   `PCTGQTRS`: Percent of tract population residing in group quarters - from USDA Food Atlas

-   `lalowihalfshare`: Share of tract population that are low income individuals beyond 1/2 mile from supermarket - from USDA Food Atlas

-   `lasnaphalfshare`: Share of tract housing units receiving SNAP benefits count beyond 0.5 mile from supermarket - from USDA Food Atlas

-   `TractSNAP`: Total count of housing units receiving SNAP benefits in tract - from USDA Food Atlas

**Possible Additional Overlay Variables:**

-    Org type / primary services

-   Services per person per month

-   Diversity score? based on the racial categories served

**Possible Additional Background Variables**

-   Race-specific food insecurity vars from Food Atlas

-   Other food services available (not sure on data source for this one....)

-   Healthy Places Index Data (haven't fully dug into this yet)

```{r}
library(readr)
library(tidyverse)
library(sf)

HFC <- read_csv("how/raw_data/HFC_how.csv")

# make each zipcode into a unique row 
# make some overlay columns (ie various ways to describe HFC recipients)
long = HFC %>%
  # Separate ZIP codes into rows
  separate_rows(`Service Locations`, sep = ",\\s*") %>%
  # Clean up whitespace just in case
  mutate(`Service Locations` = str_trim(`Service Locations`)) %>%
  # Optional: rename the column for clarity
  rename(zip = `Service Locations`) %>% 
  group_by(`Org ID`) %>% 
  mutate(AvgTotMonthly = `Total served monthly` / length(unique(zip)),
         AvgUniqueInds = `Unique individuals served monthly`/ length(unique(zip)))

```

```{r}
# Get lat/lon for zip codes 
zipcode_info <- zipcodeR::zip_code_db
zipcode_info$housing_ratio = zipcode_info$occupied_housing_units / zipcode_info$housing_units
HFC_geo <- long %>%
  left_join(zipcode_info, by = c("zip" = "zipcode")) %>% 
  filter(!is.na(lat) & !is.na(lng))
 # filter(lat >33.5)#remove weird island thing 


```

```{r map of zipcode level housing_ratio}

# Filter only ZIPs in Los Angeles County
la_zip_sf <- zipcode_info %>%
  filter(major_city == "Los Angeles" & state == "CA") %>%
  filter(!is.na(lat) & !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326, remove = FALSE)

library(tigris)
options(tigris_use_cache = TRUE)

# Get ZIP tabulation areas for CA
zcta <- zctas(cb = TRUE, year = 2020)


# 1. Compute count of each ZIP in `long`
zip_counts <- long %>%
  count(zip, name = "hfc_count") %>%
  mutate(zip = as.character(zip))


# 2. Join counts to HFC_geo (before filtering)
HFC_geo <- long %>%
  left_join(zipcode_info, by = c("zip" = "zipcode")) %>%
  filter(!is.na(lat) & !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  left_join(zip_counts, by = c("zip"))

# 3. Filter HFC_geo to valid ZIPs (same as before)
zcta_la <- zcta %>%
  rename(zipcode = ZCTA5CE20) %>%
  filter(zipcode %in% la_zip_sf$zipcode) %>%
  left_join(zipcode_info, by = "zipcode") %>%
  st_transform(crs = 4326)

HFC_geo_filtered <- HFC_geo %>%
  filter(zip %in% zcta_la$zipcode) %>%
  st_transform(crs = 4326)

# 4. Plot with dot size proportional to hfc_count
ggplot() +
  geom_sf(data = zcta_la, aes(fill = housing_ratio), color = "white", size = 0.2) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey80", name = "Housing Ratio") +
  geom_sf(data = HFC_geo_filtered, aes(size = hfc_count), color = "black", shape = 21, fill = "white", stroke = 0.5) +
  scale_size_continuous(range = c(1, 6), name = "HFC Recipient Orgs") +   # Adjust dot size range as desired
  labs(
    title = "Ratio of Occupied Housing Units to Total Housing Units",
    subtitle = "Dot size proportional to number of HFC recipient orgs per ZIP",
    caption = "Data: National Association of Realtors (zipcodeR), tigris, HFC"
  ) +
  theme_minimal() +
  theme(legend.position = "right")



```

```{r SNAP at CT level }
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(here)

options(tigris_use_cache = TRUE)

# Load Atlas data
atlas <- read_csv(here("how/inputs/ResearchAtlas.csv"))

atlas_la <- atlas %>%
  filter(State == "California", County == "Los Angeles County") %>%
  select(CensusTract, 
         LILATracts_halfAnd10,
         PCTGQTRS,
         lalowihalfshare,
         lasnaphalfshare,
         TractSNAP) %>%
  mutate(CensusTract = str_pad(as.character(CensusTract), 11, pad = "0"))

# 1. Standardize ZIP codes
long <- long %>%
  mutate(zip = str_pad(as.character(zip), 5, pad = "0"))

# 2. Count occurrences of each ZIP
zip_counts <- long %>%
  count(zip, name = "hfc_count")

# 3. Load ZCTA shapefile and filter to ZIPs in `long`
zcta <- zctas(cb = TRUE, year = 2020)

zcta_long <- zcta %>%
  rename(zip = ZCTA5CE20) %>%
  mutate(zip = str_pad(as.character(zip), 5, pad = "0")) %>%
  filter(zip %in% zip_counts$zip) %>%
  left_join(zip_counts, by = "zip") %>%
  st_transform(crs = 4326) %>%
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(lat = st_coordinates(centroid)[,2]) %>%
  filter(lat >= 33.5 & lat <= 34.5)

# Create ZIP centroid points for HFC dots
zip_centroids <- zcta_long %>%
  st_centroid() %>%
  mutate(lat = st_coordinates(.)[,2]) %>%
  filter(lat >= 33.5 & lat <= 34.5)

# 4. Load LA County census tracts and join with atlas SNAP data
la_tracts <- tracts(state = "CA", county = "Los Angeles", cb = TRUE, year = 2020) %>%
  st_transform(crs = 4326) %>%
  mutate(GEOID = str_pad(GEOID, 11, pad = "0")) %>%
  left_join(atlas_la, by = c("GEOID" = "CensusTract")) %>%
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(lat = st_coordinates(centroid)[,2]) %>%
  filter(lat >= 33.5 & lat <= 34.5)

# 5. Plot: Tract fill = TractSNAP, overlay ZIPs by count
ggplot() +
  geom_sf(data = la_tracts, aes(fill = TractSNAP), color = "white", size = 0.2) +
  geom_sf(data = zip_centroids, aes(size = hfc_count), color = "cyan",fill="cyan", alpha =0.5, shape = 21, stroke = 1) +
  scale_fill_viridis_c(option = "plasma", name = "# of SNAP Households", na.value = "grey90") +
  scale_size_continuous(range = c(1, 6), name = "# of HFC recipient orgs") +
  labs(
    title = "SNAP Participation by Census Tract (Los Angeles County)",
    subtitle = "Overlaid with number of HFC recipient orgs by ZIP code"
  ) +
  theme_minimal() +
  theme(legend.position = "right")+ 
  coord_sf(xlim = c(-118.52, -118.12), ylim = c(33.9, 34.2), expand = FALSE) #zoom in a little 


```

```{r map of HFC recipients - not in use }

library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(stringr)

options(tigris_use_cache = TRUE)

# 1. Standardize ZIP codes
long <- long %>%
  mutate(zip = str_pad(as.character(zip), 5, pad = "0"))

# 2. Count occurrences of each ZIP
zip_counts <- long %>%
  count(zip, name = "hfc_count")

# 3. Load ZCTA shapefile and filter to ZIPs in `long`
zcta <- zctas(cb = TRUE, year = 2020)

zcta_long <- zcta %>%
  rename(zip = ZCTA5CE20) %>%
  mutate(zip = str_pad(as.character(zip), 5, pad = "0")) %>%
  filter(zip %in% zip_counts$zip) %>%
  left_join(zip_counts, by = "zip") %>%
  st_transform(crs = 4326) %>%
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(lat = st_coordinates(centroid)[,2]) %>%
  filter(lat >= 33.5 & lat <= 34.5)

# 4. Load LA County census tracts
la_tracts <- tracts(state = "CA", county = "Los Angeles", cb = TRUE, year = 2020) %>%
  st_transform(crs = 4326) %>%
  mutate(centroid = st_centroid(geometry)) %>%
  mutate(lat = st_coordinates(centroid)[,2]) %>%
  filter(lat >= 33.5 & lat <= 34.5)

# 5. Plot the map
ggplot() +
  geom_sf(data = la_tracts, fill = "grey90", color = "white", size = 0.2) +
  geom_sf(data = zcta_long, aes(fill = hfc_count), color = "white", size = 0.2) +
  scale_fill_viridis_c(option = "plasma", name = "# of HFC recipients", na.value = "grey80") +
  labs(
    title = "Number of HFC Recipients (Zipcode)"
    #caption = "Data: TIGRIS, HFC ZIPs"
  ) +
  theme_minimal() +
  theme(legend.position = "right")


```

```{r - scratch }
library(here)
# data from: https://www.ers.usda.gov/data-products/food-access-research-atlas/download-the-data
atlas <- read_csv(here("how/inputs/ResearchAtlas.csv"))
atlas_la <- atlas %>%
  filter(State == "California", County == "Los Angeles County") %>%
  select(CensusTract, 
         LILATracts_halfAnd10, #Flag for low-income and low access when considering low accessibilty at 1/2 and 10 miles
         PCTGQTRS, #Percent of tract population residing in group quarters
         lalowihalfshare, #Share of tract population that are low income individuals beyond 1/2 mile from supermarket
         lasnaphalfshare, #Share of tract housing units receiving SNAP benefits count beyond 0.5 mile from supermarket
         TractSNAP #Total count of housing units receiving SNAP benefits in tract
)

# want to add the race specific vars next!!!! 

library(tigris)
library(sf)
library(dplyr)

options(tigris_use_cache = TRUE)

# Get 2020 census tract geometries for LA County
la_tracts <- tracts(state = "CA", county = "Los Angeles", year = 2020, cb = TRUE)

# Join Atlas data to tract geometries
atlas_la_clean <- atlas %>%
  filter(State == "California", County == "Los Angeles County") %>%
  select(CensusTract,
         LILATracts_halfAnd10,
         PCTGQTRS,
         lalowihalfshare,
         lasnaphalfshare,
         TractSNAP) %>%
  mutate(CensusTract = as.character(CensusTract))

la_tracts <- la_tracts %>%
  mutate(GEOID = str_pad(GEOID, 11, pad = "0"))

atlas_la_clean <- atlas_la_clean %>%
  mutate(CensusTract = str_pad(CensusTract, 11, pad = "0"))


la_tracts_atlas <- la_tracts %>%
  mutate(CensusTract = GEOID) %>%
  left_join(atlas_la_clean, by = "CensusTract")

# Count number of HFCs per ZIP
hfc_zip_counts <- long %>%
  count(zip, name = "hfc_count")

zip_centroids <- zipcode_info %>%
  filter(county == "Los Angeles County", state == "CA", !is.na(lat), !is.na(lng)) %>%
  select(zipcode, lat, lng) %>%
  distinct(zipcode, .keep_all = TRUE)

hfc_labels <- hfc_zip_counts %>%
  left_join(zip_centroids, by = c("zip" = "zipcode")) %>%
  filter(!is.na(lat) & !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326)



bbox <- st_bbox(zcta_la)  # extract bounding box from ZIP shapefile

ggplot() +
  geom_sf(data = la_tracts_atlas, aes(fill = TractSNAP), color = "white", size = 0.2) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey80", name = "% in Group Quarters") +
  geom_sf(data = hfc_labels, color = "black", shape = 21, fill = "white", size = 2, stroke = 0.5) +
  geom_sf_text(data = hfc_labels, aes(label = hfc_count), size = 3, fontface = "bold") +
 # coord_sf(xlim = c(bbox_poly["xmin"], bbox_poly["xmax"]),
#           ylim = c(bbox_poly["ymin"], bbox_poly["ymax"])) +
  labs(
    title = "Group Quarters Population by Census Tract (Los Angeles County)",
    subtitle = "Labeled by number of HFCs per ZIP code",
    caption = "Data: USDA Food Access Research Atlas, zipcodeR, HFC"
  ) +
  theme_minimal() +
  theme(legend.position = "right")




```

```{r map of LA county service planning areas - not in use}

# Load the shapefile of LA County Service Planning Areas
spa_shapefile <- st_read("how/inputs/Service_Planning_Areas_(2022).shp")



# Step 1: Convert HFC_geo to an sf object (if it is not already)
HFC_geo_sf <- st_as_sf(HFC_geo, coords = c("lng", "lat"), crs = 4326)

# Step 2: Ensure both datasets are in the same CRS
spa_map_crs <- st_crs(spa_shapefile)  # Get the CRS from the map data
HFC_geo_sf <- st_transform(HFC_geo_sf, crs = spa_map_crs)  # Transform HFC_geo to match the CRS of the map

# Step 3: Create the combined map with points
spa_map_with_scatter <- ggplot(data = spa_shapefile) +
  geom_sf(color = "black", fill = "transparent") +  # Draw the spatial map
  geom_sf(data = HFC_geo_sf, aes(geometry = geometry), color = "#2980B9", size = 3, alpha = 0.7) +  # Add points
  labs(title = "HoFoCo Providers", subtitle = "Overlayed on Los Angeles SPAs") + 
  theme_minimal() +
  theme(
    panel.grid = element_blank(),   # Remove grid lines
    axis.text = element_blank(),    # Remove axis text (numbers)
    axis.ticks = element_blank(),   # Remove axis ticks
    axis.title = element_blank()    # Remove axis titles
  )

# View the combined map
spa_map_with_scatter

```
